---
title: "HW5"
author: "Caroline Dumas"
date: "February 3, 2021"
output: html_document
---
#### **Question 1:** Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z

```{r}
x <- 1.1
a <- 2.2
b <- 3.3 
```
##### a) x^a^b
 

```{r}
z <- x^a^b
print(z)
```
##### b) (x^a)^b

```{r}
z <- (x^a)^b
print(z)
```
##### c) 3x^3 + 2x^2 + 1

```{r}
z <- 3*x^3 + 2*x^2 + 1
print(z)
```

#### **Question 2:** Using the rep and seq functions, create the following vectors

##### a) (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)

```{r}
z <- c(seq(from= 1, to = 8, by = 1), seq(from = 7, to = 1, by = -1))
print(z)
```

##### b) (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)

```{r}
z <- c(1, rep(2,2), rep(3,3), rep(4,4), rep(5,5))
print(z)
```

##### c) (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)

```{r}
z <- c(5, rep(4,2), rep(3,3), rep(2,4), rep(1,5))
print(z)
```

#### **Question 3:** Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (asin(), acos(), or atan()), convert these numbers into polar coordinates

```{r}
x <- runif(1) #random uniform number x
print(x)
y <- runif(1) # random uniform number y
print(y)
z <- c(x,y) # vector of x and y 
print(z) 
a <- sqrt(x^2 + y^2) # polar coorinates from x,y
print(a)
b <- atan(y/x) # inverse trig atan function to generate theta of polar coordinate of x,y
print(b)
c <- c(a,b) # vector of x,y polar coordinates 
print(c)
```

#### **Question 4:** Suppose that queue <- c("sheep", "fox", "owl", "ant") and that queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update the queue successively as

```{r}
queue <- c("sheep", "fox", "owl", "ant") # vector for queue
```

##### a) the serpent arrives and gets in line

```{r}
queue_1 <- c(queue, "serpent")
print(queue_1)
```

##### b) the sheep enters the ark 

```{r}
queue_2 <- queue_1[-1] # removes position 1 from queue_1
print(queue_2)
```

##### c) the donkey arrives and talks his way to the front of the line

```{r}
queue_3 <- c("donkey", queue_2) # donkey then queue_2
print(queue_3)
```

##### d) the serpent gets impatient and leaves 

```{r}
queue_4 <- queue_3[-5] #removes position 5 from queue_3
print(queue_4)
```

##### e) the owl gets bored and leaves 

```{r}
queue_5 <- queue_4[-3] #removes position 3 from queue_4
print(queue_5)
```

##### f) the aphid arrives and the ant invites him to cut in line

```{r}
queue_6 <- append(queue_5, "aphid", after = 2) # append allows adding of elements to a vector at a certain position. So adding aphid to position 2 of queue_5
print(queue_6)
```

##### g) finally, determine the position of the aphid in the line

```{r}
which(queue_6 == "aphid") # which funciton tells what indices are true, so asking which element in vector queue_6 is aphid
```

#### **Question 5:** Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7

```{r}
z <- c(seq(from = 1, to = 100, by = 1)) # create vector of 1-100
print(z)
y <- z[(z %% 2 !=0) & (z %% 3 !=0) & (z %% 7 !=0)] # don't want any of the elements that have a remainder of 0 when we divide. So 'z %% 2' gives us the remainders of each element in z when divided by 0. Then, '!=0' says to not include any of the elements with a remainder of 0. The & between each function adds the commands together. 
print(y)
```


