---
title: "HW6"
author: "Caroline Dumas"
date: "March 10, 2021"
output: html_document
---
### 1. Assign to the variable n_dims a single random integer between 3 and 10.
```{r}
x <- 3:10 # vector from 3-10
n_dims <- sample(x=x, size=1) #randomly chooses one of the numbers in the vector x
print(n_dims)
```
Create a vector of consecutive integers from 1 to n_dims2
```{r}
y <- c(seq(from=1, to=n_dims^2)) # sequence to create vector 
print(y)
```
Use the sample function to randomly reshuffle these values.
```{r}
z <- sample(y) # sample randomly rearranges vector 
print(z)
```
Create a square matrix with these elements.
```{r}
z_matrix <- matrix(data=z, nrow=n_dims) # using n_dims as the nrow makes the matrix be the length of whatever the random number of n_dims 
print(z_matrix)
```
Find a function in r to transpose the matrix.
```{r}
z_transpose <- t(z_matrix)
print(z_transpose)
```
Calculate the sum and the mean of the elements in the first row and the last row
```{r}
firstRow <- z_transpose[1, ] # atomic vector of the first row
lastRow <- z_transpose[n_dims, ] # atomic vector of the last row -- n_dims is the random number chosen
firstRow + lastRow # sum of the elements in the first and last row 
sum(firstRow) # sum of elements in the first row 
sum(lastRow) # sum of the elements in the last row 
mean(firstRow) # mean of the elements in the first row 
mean(lastRow) # mean of the elements in the last row 
```
Read about the eigen() function and use it on your matrix
```{r}
?eigen
eigen_z_transpose <- eigen(z_transpose, only.values=FALSE)
print(eigen_z_transpose)
# the $values and $vectors are imaginary numbers 
typeof(eigen_z_transpose)
typeof(eigen_z_transpose$values)
typeof(eigen_z_transpose$vectors)
```
If have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.
```{r}
# use n_dims for row length or column length so that it will always work 
```
### Create a list with the following named elements:
*my_matrix*, which is a 4 x 4 matrix filled with random uniform values
```{r}
my_matrix <- matrix(data=runif(16), nrow=4)
print(my_matrix)
```
*my_logical* which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
```{r}
my_logical <- runif(100) > 0.5 # inequality sign makes it a logical 
head(my_logical)
```
*my_letters*, which is a 26-element vector of all the lower-case letters in random order.
```{r}
my_letters <- sample(letters[1:26], replace=FALSE) # randomly sample letters from 1-26 without repleating any
print(my_letters)
```
create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r}
list <- list(my_matrix[[2,2]], my_logical[[2]], my_letters[[2]])
print(list)
```
use the typeof() function to confirm the underlying data types of each component in this list
```{r}
typeof(list[[1]])
typeof(list[[2]])
typeof(list[[3]])
```
combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}
vector_list <- c(list[[1]], list[[2]], list[[3]])
print(vector_list)
```
what is the data type of this vector?
```{r}
typeof(vector_list) # everyone coerced to a character
```
### Create a data frame with two variables (= columns) and 26 cases (= rows).
call the first variable my_unis and fill it with 26 random uniform values from 0 to 10. call the second variable my_letters and fill it with 26 capital letters in random order.
```{r}
my_unis <- runif(26, min=0, max=10)
print(my_unis)
my_letters <- sample(LETTERS)
print(my_letters)
d_frame <- data.frame(my_unis, my_letters)
head(d_frame)
```
for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.
```{r}
d_frame[sample(my_unis, size=4), 1] <- NA # sample 4 items in my_unis then specify only in the first column 
print(d_frame)
```
for the first variable, write a single line of R code to identify which rows have the missing values.
```{r}
which(!complete.cases(d_frame)) # asking which items are not complete cases (NA)
```
for the second variable, sort it in alphabetical order
```{r}
d_frame[order(d_frame$my_letters),] # order function to order the letters
```
calculate the column mean for the first variable.
```{r}
mean(d_frame$my_unis, na.rm=TRUE) # have to na.rm=TRUE to remove the NA. mean function default is to keep them.
```

